---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# noah

<!-- badges: start -->

[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://www.tidyverse.org/lifecycle/#experimental) [![CRAN status](https://www.r-pkg.org/badges/version/noah)](https://CRAN.R-project.org/package=noah) [![R build status](https://github.com/Teebusch/noah/workflows/R-CMD-check/badge.svg)](https://github.com/Teebusch/noah/actions) [![Codecov test coverage](https://codecov.io/gh/Teebusch/noah/branch/master/graph/badge.svg)](https://codecov.io/gh/Teebusch/noah?branch=master)

[![R build status](https://github.com/Teebusch/noah/workflows/R-CMD-check/badge.svg)](https://github.com/Teebusch/noah/actions)
<!-- badges: end -->

`noah` (**no** **a**nimals were **h**armed) generates pseudonyms that are 
delightful and easy to remember. 
Instead of cryptic alphanumeric IDs, `noah` generates anonymous animals like the
*Likeable Leech* and the *Proud Chickadee*.

## Installation

You can install the development version of noah from [Github](/https://github.com/teebusch/noah) with:

```{r eval=FALSE}
# install.packages("remotes")
remotes::install_github("teebusch/noah")
```

## Usage

```{r include=FALSE}
set.seed(21122020)
```

### Generating pseudonyms with `noah`

A quick way to create pseudonyms with `noah` is to use the `pseudonymize()`
function. It will generate pseudonyms for every element in a vector:

```{r}
library(noah)

pseudonymize(1:6)
```

Repeated elements will receive the same pseudonym:

```{r}
pseudonymize(rep(1:2, each = 2))
```

`pseudonymize()` takes any number of input vectors, as long as they are the same 
length. It will treat the elements in the same position as being from the same 
subject. 

```{r}
pseudonymize(
  c("😙", "😙", "😙"), 
  c("🥕", "🥕", "🍰")
)
```

### Adding pseudonyms to a data frame

Often we may want to add a column with pseudonyms to a data frame, using one or 
more of the columns as identifiers. We can do this with `pseudonymize()` and
`dplyr::mutate()`.

Here we use the diabetic retinopathy dataset from the `survival` 
package and add a new column with a pseudonym for each unique `id`:

```{r example-pseudonymize, warning=FALSE, message=FALSE}
library(dplyr)
diabetic <- dplyr::as_tibble(survival::diabetic)

diabetic %>% 
  mutate(pseudonym = pseudonymize(id)) %>% 
  relocate(pseudonym)
```


# Keeping track of pseudonyms

Internally, `pseudonymize()` uses an object of class `Ark` that acts like a 
pseudonym dictionary and keeps track of pseudonyms that have been used.
Normally, a new `Ark` is built for each call of the `pseudonymize()` function.
However, we can use an `Ark` to ensure that the same input is always assigned 
the same pseudonym across multiple data sets:

```{r example-ark}
ark <- Ark$new()

# split dataset into left and right eye and pseudonymize separately
diabetic_left <- diabetic %>% 
  filter(eye == "left") %>% 
  mutate(pseudonym = pseudonymize(id, .ark = ark))

diabetic_right <- diabetic %>% 
  filter(eye == "right") %>% 
  mutate(pseudonym = pseudonymize(id, .ark = ark))

# reunite the data sets again
bind_rows(diabetic_left, diabetic_right) %>% 
  relocate(pseudonym) %>% 
  arrange(id)
```
