---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# noah

<!-- badges: start -->

[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://www.tidyverse.org/lifecycle/#experimental) [![CRAN status](https://www.r-pkg.org/badges/version/noah)](https://CRAN.R-project.org/package=noah) [![R build status](https://github.com/Teebusch/noah/workflows/R-CMD-check/badge.svg)](https://github.com/Teebusch/noah/actions) [![Codecov test coverage](https://codecov.io/gh/Teebusch/noah/branch/master/graph/badge.svg)](https://codecov.io/gh/Teebusch/noah?branch=master)

<!-- badges: end -->

`noah` (**no** **a**nimals were **h**armed) generates pseudonyms that are 
delightful and easy to remember. 
Instead of cryptic alphanumeric IDs, it creates anonymous animals like the
*Likeable Leech* and the *Proud Chickadee*.

## Installation

You can install `noah` from [Github](/https://github.com/teebusch/noah) with:

```{r eval=FALSE}
# install.packages("remotes")
remotes::install_github("teebusch/noah")
```

## Usage

```{r include=FALSE}
set.seed(122020)
```

### Generating pseudonyms

A quick way to create pseudonyms with `noah` is the `pseudonymize()`
function. It will generate pseudonyms for every element in a vector, so if you 
need 100 pseudonyms, simply call `pseudonymize(1:100)`. 
Note that repeated elements will receive the same pseudonym:

```{r}
library(noah)

pseudonymize(rep(1:4, times = 2))
```

`pseudonymize()` accepts any number of input vectors as arguments, as long as 
they have the same length. It will treat elements in the same position as 
being from the same subject. 

```{r}
pseudonymize(c("🐰", "🐰", "🐰"), c("🥕", "🥕", "🍰"))
```

### Adding pseudonyms to a data frame

Often we want to add a column with pseudonyms to a data frame, using one or 
more columns as identifiers. We can do this with `pseudonymize()` and
`dplyr::mutate()`.

In this example we use the diabetic retinopathy dataset from the `survival` 
package and add a new column with a pseudonym for each unique `id`:

```{r example-pseudonymize, warning=FALSE, message=FALSE}
library(dplyr)
diabetic <- as_tibble(survival::diabetic)

diabetic %>% 
  mutate(pseudonym = pseudonymize(id)) %>% 
  relocate(pseudonym)
```

### Keeping track of pseudonyms

Internally, `pseudonymize()` uses an object of class `Ark` that keeps track of 
which pseudonyms that have been used.
Normally, a new `Ark` is created for each call of the `pseudonymize()` function,
but we can use an `Ark` to ensure that the same input is always assigned 
the same pseudonym across multiple data sets:

```{r example-ark}
ark <- Ark$new()

# split dataset into left and right eye and pseudonymize separately
diabetic_left <- diabetic %>% 
  filter(eye == "left") %>% 
  mutate(pseudonym = pseudonymize(id, .ark = ark))

diabetic_right <- diabetic %>% 
  filter(eye == "right") %>% 
  mutate(pseudonym = pseudonymize(id, .ark = ark))

# reunite the data sets again
bind_rows(diabetic_left, diabetic_right) %>% 
  relocate(pseudonym) %>% 
  arrange(id)
```

## Related R packages

There are multiple R packages that generate fake data, including names,
phone numbers, addresses, credit card numbers, gene sequences and more:

- [`charlatan`](https://docs.ropensci.org/charlatan/)
- [`randomNames`](https://centerforassessment.github.io/randomNames/)
- [`randNames`](https://github.com/karthik/randNames)
- [`generator`](https://github.com/paulhendricks/generator).
  
There are also packages that can anonymize personal identifiable information.
If you need reliable, watertight anonymization, `noah` is not the right tool for
you. Instead, check out these packages:

- [`sdcMicro`](http://sdctools.github.io/sdcMicro/index.html)
- [`sdcTable`](https://sdctools.github.io/sdcTable/index.html)
- [`anonymizer`](http://paulhendricks.io/anonymizer/)
- [`digest`](http://dirk.eddelbuettel.com/code/digest.html)

