---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# noah

<!-- badges: start -->

[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://www.tidyverse.org/lifecycle/#experimental) [![CRAN status](https://www.r-pkg.org/badges/version/noah)](https://CRAN.R-project.org/package=noah) [![R build status](https://github.com/Teebusch/noah/workflows/R-CMD-check/badge.svg)](https://github.com/Teebusch/noah/actions) [![Codecov test coverage](https://codecov.io/gh/Teebusch/noah/branch/master/graph/badge.svg)](https://codecov.io/gh/Teebusch/noah?branch=master)

<!-- badges: end -->

noah (**no** **a**nimals were **h**armed) generates pseudonyms that are 
delightful and easy to remember. 
Instead of cryptic alphanumeric IDs, it creates anonymous animals like the
*Likeable Leech* and the *Proud Chickadee*.

## Installation

You can install noah from [Github](/https://github.com/teebusch/noah) with:

```{r eval=FALSE}
# install.packages("remotes")
remotes::install_github("teebusch/noah")
```

## Usage

```{r include=FALSE}
set.seed(122020)
```

### Generating pseudonyms

The `pseudonymize()` function generates pseudonyms for every element 
in a vector or every row in a data frame, whereby repeated elements receive the
same pseudonym.
If you need 100 pseudonyms, call `pseudonymize(1:100)`. 

```{r}
library(noah)

pseudonymize(rep(1:4, times = 2))
```

`pseudonymize()` accepts any number of input vectors as arguments, as long as 
they have the same length. It will treat elements in the same position as 
being from the same subject. 

```{r}
pseudonymize(c("🐰", "🐰", "🐰"), c("🥕", "🥕", "🍰"))
```

### Adding pseudonyms to data frames

Often we want to add a column with pseudonyms to a data frame, using one or 
more columns as identifiers. 
In this example we use the diabetic retinopathy dataset from the `{survival}` 
package.
There are two was to do add pseudonyms to a data frame with noah:

#### Using `mutate()` and `pseudonymize()`

We can add pseudonyms using `pseudonymize()` and `dplyr::mutate()`. 
Here we add a new column with a pseudonym for each unique id:

```{r warning=FALSE, message=FALSE}
library(dplyr)
diabetic <- as_tibble(survival::diabetic)

diabetic %>% 
  mutate(pseudonym = pseudonymize(id)) %>% 
  relocate(pseudonym)
```

#### Using `add_pseudonyms()`

We can also use the `add_pseudonyms()` function, which wraps the mutate and 
relocate step and supports `tidyselect` syntax for the selection of the 
identifier columns:

```{r}
diabetic %>% 
  add_pseudonyms(where(is.factor))
```

### Keeping track of pseudonyms

Internally, `pseudonymize()` and `add_pseudonyms()` use an object of class `Ark`
(a pseudonym archive) to keeps track of the pseudonyms that have been used.
By default, a new `Ark` is created for each call of the `pseudonymize()` 
function, but we can provide an `Ark` to ensure that the same input is always
assigned the same pseudonym across multiple function calls:

```{r example-ark}
ark <- Ark$new()
```

```{r echo=FALSE}
ark
```

```{r}
# split dataset into left and right eye and pseudonymize separately
diabetic_left <- diabetic %>% 
  filter(eye == "left") %>% 
  add_pseudonyms(id, .ark = ark)

diabetic_right <- diabetic %>% 
  filter(eye == "right") %>% 
  add_pseudonyms(id, .ark = ark)

# reunite the data sets again
bind_rows(diabetic_left, diabetic_right) %>% 
  arrange(id)
```

If we have a look at the ark now, we can see that it contains `r length(ark)`
pseudonyms -- just as many as there are unique id's in the dataset.

```{r echo=FALSE}
ark
```


## Related R packages

There are multiple R packages that generate fake data, including fake names,
phone numbers, addresses, credit card numbers, gene sequences and more:

- [`{charlatan}`](https://docs.ropensci.org/charlatan/)
- [`{randomNames}`](https://centerforassessment.github.io/randomNames/)
- [`{randNames}`](https://github.com/karthik/randNames)
- [`{generator}`](https://github.com/paulhendricks/generator).
  
There are also packages for anonymizing personal identifiable information in data
sets. 
If you need reliable, watertight anonymization, noah is likely not the right
tool for you and you should check out these packages instead:

- [`{sdcMicro}`](http://sdctools.github.io/sdcMicro/index.html)
- [`{sdcTable}`](https://sdctools.github.io/sdcTable/index.html)
- [`{anonymizer}`](http://paulhendricks.io/anonymizer/)

